% Place your vita below.

\textbf{Education}
\begin{itemize}
    \item Doctor of Philosophy, Physics, The Pennsylvania State University, University Park, Pennsylvania, USA, 2023
    \item Bachelor of Science, Physics, The University of Minnesota, Minneapolis, Minnesota, USA, 2017
\end{itemize}

\textbf{Selected Publications}
\begin{itemize}
    \item Astari Esfahani, A. et al. (2023) "Antenna Arrays for CRES-based Neutrino Mass Measurement", \emph{Phys. Rev. C}, In preparation.
    \item Astari Esfahani, A. et al. (2023) "Real-time Signal Detection for Cyclotron Radiation Emission Spectroscopy Measurements using Antenna Arrays", \emph{Journal of Instrumentation}, In preparation.
    \item Astari Esfahani, A. et al. (2023) "Tritium Beta Spectrum and Neutrino Mass Limit from cyclotron Radiation Emission Spectroscopy", \emph{Phys. Rev. Lett.}, Accepted.
    \item Astari Esfahani, A. et al. (2022) "SYNCA: A Synthetic Cyclotron Antenna for the Project 8 Collaboration", \emph{Journal of Instrumentation}, \textbf{18}(01).
\end{itemize}

\textbf{Selected Presentations}
\begin{itemize}
    \item \emph{New Developments in the CRES Technique for Neutrino Mass Measurement}, Invited Talk, Fall 2022 Meeting of the APS Division of Nuclear Physics, New Orleans, Louisiana, USA, 2022
    \item \emph{Signal Detection Algorithms for Phase III of the Project 8 Experiment}, Contributed Talk, APS April Meeting 2022, New York, New York, USA, 2022
    \item \emph{Synthetic Electron Antenna for Calibrating the Project 8 Neutrino Mass Experiment}, Contributed Talk, Fall 2021 Meeting of the APS Division of Nuclear Physics, Virtual, 2021
\end{itemize}

